Covariates analysis of MSBB RNASeq data of region BM 44 (realigned from source FASTQ with STAR-HTSeq)

### 2016-09-10 code/Rmd/MSSM_BM_44_RNASeq_Reprocessed_LillyPipeline.Rmd
Added markdown to process BM 44 in MSSM (reprocessed by eli lilly)

knit2synapse::knitToFolderEntity(file = "./MSSM_BM_44_RNASeq_Realigned_StarHTSeq.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB BM 44 RNASeq RealignedFromSource StarHTSeq')
```

@@

thisFileName <- 'MSSM_BM_44_RNASeq_Realigned_StarHTSeq.Rmd'

@@

Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
COUNT_ID <- 'syn7221166'
ALL_USED_IDs <- COUNT_ID
COUNT <- read.table(synGet(COUNT_ID)@filePath, quote = '', header = T, sep = '\t')

# Get sample ids
SampleID = data.frame(SampleID = colnames(COUNT),
                      ID = colnames(COUNT)) %>% 
  tidyr::separate(ID, c('A','B','ID', 'C'), sep = '_')

@@

METADATA = left_join(METADATA_TECHNICAL %>%
                       dplyr::mutate(ID = gsub('hB_RNA_','',NewBarcode),
                                     ID = as.numeric(ID)),
                     METADATA_CLINICAL %>%
                       dplyr::select(BB, BM.44.Barcode, PMI, RACE, 
                                     AOD, CDR, SEX, NP.1, PlaqueMean, bbscore) %>%
                       dplyr::rename(BM_44 = BM.44.Barcode) %>%
                       tidyr::gather(BrainRegion, ID, -BB, -PMI, -RACE, 
                                     -AOD, -CDR, -SEX, -NP.1, -PlaqueMean, -bbscore) %>%
                       filter(!is.na(ID)) %>%
                       dplyr::mutate(ID = as.numeric(ID))) %>%
  inner_join(SampleID) %>%
  filter(!is.na(SampleID)) %>%
  unique
rownames(METADATA) = METADATA$SampleID

@@

### Outlier analysis
Clustering of initial normalised data (with NULL design)
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E, scale.=T)

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=CDR, shape=SEX, size=PMI))
p <- p + theme_bw() + theme(legend.position="top") + scale_shape_manual(values = 1:7)
# p <- p + geom_text(aes(label= ID), size=4, hjust=1)
p

@@

```{r residualSigCovars.manual, fig.width=20, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "CDR" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = c("CDR0","CDR0.5","CDR1","CDR2","CDR3","CDR4","CDR5")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```
Coexpression of genes
```{r coexp2}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
### Clustering residual data
```{r decompse.normalise.data2, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T)

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = "MSBB BM 44 RNASeq  RealignedFromSource StarHTSeq", parentId = parentId)
CODE <- synStore(CODE)

>>>>

knit2synapse::knitToFolderEntity(file = "./MSSM_BM_44_RNASeq_Reprocessed_LillyPipeline.Rmd",
                                 parentId ="syn5570248",
                                 entityName = 'MSBB BM 44 RNASeq Reprocessed LillyPipeline')
```

@@

thisFileName <- 'MSSM_BM_44_RNASeq_Reprocessed_LillyPipeline.Rmd'

@@

Obtain count matrix and metadata from synapse.
```{r download.data, cache=TRUE}
# Download expression data
COUNT_B1_ID <- 'syn6114414'
ALL_USED_IDs <- COUNT_B1_ID
COUNT_B1 <- read.table(synGet(COUNT_B1_ID)@filePath, quote = '', header = T, sep = '\t')

COUNT_B2_3_ID <- 'syn6125203'
ALL_USED_IDs[length(ALL_USED_IDs)+1] <- COUNT_B2_3_ID
COUNT_B2_3 <- read.table(synGet(COUNT_B2_3_ID)@filePath, quote = '', header = T, sep = '\t')

# Download expression data
COUNT_B4_ID <- 'syn6130073'
ALL_USED_IDs[length(ALL_USED_IDs)+1] <- COUNT_B4_ID
COUNT_B4 <- read.table(synGet(COUNT_B4_ID)@filePath, quote = '', header = T, sep = '\t')

# Remove samples in Batch 2&3 if they present in batch 4
COUNT_B2_3 = COUNT_B2_3[,setdiff(colnames(COUNT_B2_3), colnames(COUNT_B4))]

COUNT = cbind(COUNT_B1, COUNT_B2_3[rownames(COUNT_B1),], COUNT_B4[rownames(COUNT_B1),])

# Get sample ids
SampleID = data.frame(SampleID = colnames(COUNT),
                 ID = colnames(COUNT)) %>% 
  tidyr::separate(ID, c('A','B','ID'), sep = '_') 
SampleID$ID[SampleID$ID == 'HISEQ'] = SampleID$B[SampleID$ID == 'HISEQ']

@@

METADATA = left_join(METADATA_TECHNICAL %>%
                       dplyr::mutate(ID = gsub('hB_RNA_','',NewBarcode),
                                     ID = as.numeric(ID)),
                     METADATA_CLINICAL %>%
                       dplyr::select(BB, BM.44.Barcode, BM36.Fnl, BM22.Fnl, BM10.Fnl, PMI, RACE, 
                                     AOD, CDR, SEX, NP.1, PlaqueMean, bbscore) %>%
                       dplyr::rename(BM_44 = BM.44.Barcode,
                                     BM_10 = BM10.Fnl,
                                     BM_22 = BM22.Fnl,
                                     BM_36 = BM36.Fnl) %>%
                       tidyr::gather(BrainRegion, ID, -BB, -PMI, -RACE, 
                                     -AOD, -CDR, -SEX, -NP.1, -PlaqueMean, -bbscore) %>%
                       filter(!is.na(ID)) %>%
                       dplyr::mutate(ID = as.numeric(ID))) %>%
  inner_join(SampleID) %>%
  filter(!is.na(SampleID)) %>%
  unique %>%
  filter(BrainRegion == 'BM_44', !(SampleID %in% c('hB_RNA_8565', 'hB_RNA_4919', 'hB_RNA_9005')))
rownames(METADATA) = METADATA$SampleID

@@

### Outlier analysis
Clustering of initial normalised data (with NULL design)
```{r decompse.normalise.data, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(VOOM.GENE_EXPRESSION$E, scale.=T, center = T)

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(color=CDR, shape=SEX, size=PMI))
p <- p + theme_bw() + theme(legend.position="top") + scale_shape_manual(values = 1:7)
p <- p + geom_text(aes(label= ID), size=4, hjust=0)
p

@@

```{r residualSigCovars.manual, fig.width=20, fig.height=12}
residualSigCovars[["PC_res"]][[2]]$plotData
```
### Residual calculation
Calculate weighted residuals and add back "CDR" to the residuals
```{r varsToAddBack}
# Add variable of interest back to the residuals
varsToAddIn = c("CDR0","CDR0.5","CDR1","CDR2","CDR3","CDR4","CDR5")
RESIDUAL.GENE_EXPRESSION = RESIDUAL.GENE_EXPRESSION + 
  VOOM.ADJUSTED.FIT$coefficients[,varsToAddIn] %*% t(DM1$design[,varsToAddIn])
```
Coexpression of genes
```{r coexp2}
cr = cor(t(RESIDUAL.GENE_EXPRESSION))
hist(cr, main = 'Distribution of correlation between genes', xlab = 'Correlation')
```
### Clustering residual data
```{r decompse.normalise.data2, fig.height=8, fig.width=8, results='asis'}
# Find principal components of expression to plot
PC <- prcomp(RESIDUAL.GENE_EXPRESSION, scale.=T, center = T)

@@

### Store files in synapse
```{r synapse.store, include=FALSE, eval=TRUE, cache=FALSE}
# Code
CODE <- Folder(name = 'MSBB BM 44 RNASeq Reprocessed LillyPipeline', parentId = parentId)
CODE <- synStore(CODE)
